Discrete Quantum Theories and Computing

### 2019-02-06 dissertation.tex
dissertation.lyx Fix space and place of floats...

%% LyX 2.3.2-2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{rotfloat}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

@@

@@

\addlinespace
Committee or course & Gerardo Ortiz & Amr\nolinebreak[3] A. Sabry\linebreak[3] (\textarabic{عمرو صبري}) & Andrew\nolinebreak[3] J. Hanson & Dylan\nolinebreak[3] Paul Thurston & Shouhong Wang\linebreak[3] (汪守宏) & Lawrence\nolinebreak[3] S.\linebreak[2] Moss & Relation with the dissertation\tabularnewline\addlinespace

@@

\addlinespace

@@

is based on~\cite{THOS2018}.\tabularnewline\addlinespace

@@

some of the remaining ones are explained in Chapter~\ref{chap:Further-Discussion}.\tabularnewline\addlinespace

@@

inspires Chapter~\ref{chap:Introduction}.\tabularnewline\addlinespace

@@

exam (chaired by Prof.\ Sabry) inspires Chapter~\ref{chap:Introduction}.\tabularnewline\addlinespace

@@

\ref{chap:DQT=000026DQC}, and~\ref{chap:QIVPM}.\tabularnewline\addlinespace

@@

is cited in Sec.~\ref{discretequantumtheoryIIb}.\tabularnewline\addlinespace

@@

are cited in Sec.~\ref{modalquantum}.\tabularnewline\addlinespace

@@

cited in Sec.~\ref{subsec:Classical-and-Quantum}.\tabularnewline\addlinespace

@@

@@

To have the current the content of the dissertation, I would also
like to thank John Gardiner for some inspiring discussion about quantum
probability measures over finite fields~\cite{Gardiner2014}. His
results were improved and motivated the transition from DQT to QIVPM
in Sec.~\ref{sec:Toward-IVPM}. Besides of my committee members,
the knowledge I learned from teachers in other courses help me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Although
not directly relating to the dissertation, I have to thank for your
teaching to make me more mature and easier to communicate with others,
especially my English teachers, Traci Nagle, Elizabeth ``Betsy''
Merceron, and Kexin ``Casey'' Chen (谌可心).
\begin{table}[t]
\noindent \begin{centering}
\caption{\label{tab:teacher}The knowledge I learned from teachers in other
courses help me consolidate this dissertation, where IUB stands for
Indiana University Bloomington and NTU stands for National Taiwan
University.}
\par\end{centering}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{Jin-Tzu~Chen}cM{MATH-M522}Y}
\toprule 
\addlinespace
Instructor & School & Course & Relation with the dissertation\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
Tom Lewis & IUB & SLST-T501 & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper of \emph{Academic Writing}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Michael\nolinebreak[3] A. Mandell & IUB & MATH-M522 & The textbook~\cite{Hatcher2001} of \emph{Topology II} is cited
in Chapters~\ref{chap:Introduction} and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Steven Myers & IUB & CSCI-B502 & The textbook~\cite{AroraBarak2009} of \emph{Computational Complexity}
is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Valery Lunts & IUB & MATH-M501 & The textbook~\cite{DummitFoote2004} of \emph{Survey of Algebra}
is cited in Sec.~\ref{sec:background}, \ref{discretequantumtheoryI},
and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline\addlinespace
\midrule 
\addlinespace
Huah Chu\linebreak[3] (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & The textbook~\cite{Artin1991} of \emph{Algebra (I)} and \emph{(II)}
is cited in Sec.~\ref{CQC1qubitBloch.sec}, \ref{sec:background},
\ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline\addlinespace
\midrule
\addlinespace
Jin-Tzu Chen\linebreak[3] (陳金次) & NTU & 201~31300 & The textbook~\cite{GAMELIN2003} of \emph{Functions of a Complex Variable}
is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{table}

@@

\begin{sidewaystable}
\caption{\label{tab:Comparison}Comparison between the classical quantities
and their mathematical models with the inspired quantum models.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\midrule 
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\midrule 
\addlinespace
Computational model & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
How does the theoretical model predict the behavior of the physical
device? & Reliably & Not reliably, but after we have better technology to manipulate the
minimum units, we can more reliably predict the physical phenomenon
and their computational power by a better discrete model. & Not reliably, because precision can never be high enough, and the
difference between their computational power might not be able compensated
by error analysis techniques.\tabularnewline\addlinespace
\midrule 
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states, which are assumed in the field of complex numbers in the ``conventional
quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}. If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probability'' has well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\begin{comment}
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }
\end{comment}
This idea will be extended to quantum domain as quantum interval-valued
probability measure (QIVPM).

@@

\begin{sidewaystable}
\caption{\label{tab:organization}Comparison among our models, where the upper
half of the table lists their quantum mechanical definitions and properties,
and the lower half of the table lists their computational power.}

%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\midrule 
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\midrule 
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\midrule 
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\midrule 
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\midrule 
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and finally, Chapter~\ref{chap:Further-Discussion} contains further
extended questions, especially, briefly explaining why it might be
hard to recover Gleason's theorem for general QIVPMs.

@@

We can reduce 3 degrees of freedom in \Eq{A5.eq} to 2 degrees of
freedom by effectively removing $\rme^{\rmi\theta}$ (``fibering
out by the circle $\Sphere{1}$''). The standard form of these maps
(``the Hopf fibration'') is 
\begin{eqnarray}
 &  & X=2\,\mathrm{Re}\ \alpha_{0}\alpha_{1}^{*}=2x_{0}x_{1}+2y_{0}y_{1}\,,\nonumber \\
 &  & Y=2\,\mathrm{Im}\ \alpha_{0}\alpha_{1}^{*}=2x_{1}y_{0}-2x_{0}y_{1}\,,\label{eq:HopfFibration}\\
 &  & Z=\left|\alpha_{0}\right|^{2}-\left|\alpha_{1}\right|^{2}={x_{0}}^{2}+{y_{0}}^{2}-{x_{1}}^{2}-{y_{1}}^{2}\,.\nonumber 
\end{eqnarray}
By denoting the three-dimensional vector~$\left(X,Y,Z\right)$ as
$\Hat{a}$, \Eq{A5.eq} implies these transformed coordinates obeying
\begin{equation}
\left\Vert \Hat{a}\right\Vert ^{2}=X^{2}+Y^{2}+Z^{2}=\left(\left|\alpha_{0}\right|^{2}+\left|\alpha_{1}\right|^{2}\right)^{2}=1\label{A7.eq}
\end{equation}
and therefore have only two remaining degrees of freedom describing
all possible distinct one-qubit quantum states. In Fig.~\ref{hopfFibers.fig},
we illustrate schematically the family of circles \textit{each one
of which is collapsed to a point}~$\left(\phi,\psi\right)$ on the
surface $X^{2}+Y^{2}+Z^{2}=1$ by the Hopf map.

\begin{figure}[tb]
\hfill{}\subfloat[]{\centering{}\includegraphics[width=0.48\columnwidth]{hopfS2}}\hfill{}\subfloat[]{\begin{centering}
\includegraphics[width=0.48\columnwidth]{hopfFibers}
\par\end{centering}
}\hfill{}\caption{\label{hopfFibers.fig}(a) The two-sphere~$\Sphere{2}$ represented
by \Eq{A7.eq}, which is the irreducible space of one-qubit states,
along with a representative set of points on the sphere. Each single
point on the sphere in (a) corresponding to a circle in (b), and a
whole family of circles (the paths of $\rme^{\rmi\theta}$) on the
three-sphere~$\Sphere{3}$ represents the Hopf fibration, Eq.~(\ref{eq:HopfFibration}).
Although $\Sphere{3}$ cannot be directly embedded in $\R^{3}$, three-sphere~$\Sphere{3}$
can be regarded as attaching two three-dimensional balls on two sides
of two-sphere~$\Sphere{2}$. In this way, each circle in $\Sphere{3}$
can be represented as a circle in the three-dimensional ball as shown
in (b). Moreover, points in (a) are color-coded corresponding to circles
in (b), e.g., one pole contains the red elliptical circle that would
become an infinite-radius circle by a slightly different way to represent
$\Sphere{3}$ in $\R^{3}$, and the opposite pole corresponds to the
large perfectly round red circle at the equator.}
\end{figure}

The resulting manifold is the two-sphere $\Sphere{2}$ (a 2-manifold)
embedded in $\R^{3}$. If we choose one of many possible coordinate
systems describing $\Sphere{3}$ via \Eq{A5.eq} such as 
\begin{equation}
\left(x_{0},y_{0},x_{1},y_{1}\right)=\left(\cos\left(\theta+\phi\right)\cos\psi,\,\sin\left(\theta+\phi\right)\cos\psi,\,\cos\left(\theta-\phi\right)\sin\psi,\,\sin\left(\theta-\phi\right)\sin\psi\right)\,,\label{A8a.eq}
\end{equation}
where %% $0\leq \theta \leq 2 \pi$,  $-\pi \leq \phi \leq  \pi$, 
$0\leq\psi\leq\frac{\pi}{2}$, with $0\leq\theta+\phi<2\pi$ and $0\leq\theta-\phi<2\pi$,
we see that 
\begin{equation}
\left(X,Y,Z\right)=\left(\cos\left(2\phi\right)\sin\left(2\psi\right),\,\sin\left(2\phi\right)\sin\left(2\psi\right),\,\cos\left(2\psi\right)\right)\,.\label{A8b.eq}
\end{equation}
Thus, the one-qubit state is independent of $\theta$, and we can
choose $\theta=\phi$ without loss of generality, reducing the formula
of the unique one-qubit states to $\ket{\psi_{1}}=\rme^{2\rmi\phi}\cos\psi\ket{0}+\sin\psi\ket{1}$,
and an irreducible state can be represented as a point on a sphere
called the Bloch sphere, as shown in Fig.~\ref{blochSphere.fig}~(a).

\begin{figure}[tb]
\begin{centering}
\hfill{}\subfloat[]{\centering{}\includegraphics[width=0.47\columnwidth]{blochSphere}}\hfill{}\subfloat[]{\begin{centering}
\includegraphics[width=0.47\columnwidth]{blochSphereArc}
\par\end{centering}
}\hfill{}
\par\end{centering}
\caption{\label{blochSphere.fig}(a) The conventional Bloch sphere with a unique
state represented by the point at the red sphere. (b) The geodesic
shortest-distance arc connecting two one-qubit quantum states.}
\end{figure}

@@

We can now compute the size of the equivalence class of complex unit-modulus
phases corresponding to the Hopf fibration circle. %
\begin{comment}
\todo{Explain the Hopf fibration circle here\ldots ? }
\end{comment}
Since $\Fpp$ has $p^{2}-1$ non-zero values, and the map $\fnorm{\alpha}$
distributes these equally across the domain of $p-1$ non-zero elements
$c\in\Fp$, there are $\frac{p^{2}-1}{p-1}=p+1$ (non-zero) domain
elements in $\Fpp$ for each (non-zero) image element in $\Fp$. We
illustrate this graphically in Figure~\ref{proof1.fig}. Thus, the
Hopf circle always has the size $p+1$, corresponding essentially
to a discrete projective line, and that is the size of each equivalence
class of the map $\fnorm{\alpha}$ for non-vanishing $\alpha$, including
the map to the unit norm value $c=1\in\Fp$.
\begin{figure}
\noindent \centering{}\includegraphics[width=3.6in]{Fpp2Fp}\caption{\label{proof1.fig}Sketch of the map from $\Fpp$ to $\Fp$ using
$\fnorm{\alpha}$, showing the decomposition of $\Fpp$ into the zero
element $\left(0,0\right)$ and the $p^{2}-1=\left(p+1\right)\left(p-1\right)$
non-zero elements that map onto the $p-1$ non-zero elements of $\Fp$
with multiplicity $p+1$.}
\end{figure}
\end{proof}

>>>>

%% LyX 2.3.2-2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{rotfloat}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[all]{xy}

@@

\usepackage[font=doublespacing]{caption}

@@

\begin{doublespace}

@@

Committee or course & Gerardo Ortiz & Amr\nolinebreak[3] A. Sabry\linebreak[3] (\textarabic{عمرو صبري}) & Andrew\nolinebreak[3] J. Hanson & Dylan\nolinebreak[3] Paul Thurston & Shouhong Wang\linebreak[3] (汪守宏) & Lawrence\nolinebreak[3] S.\linebreak[2] Moss & Relation with the dissertation\tabularnewline

@@

@@

is based on~\cite{THOS2018}.\tabularnewline

@@

some of the remaining ones are explained in Chapter~\ref{chap:Further-Discussion}.\tabularnewline

@@

inspires Chapter~\ref{chap:Introduction}.\tabularnewline

@@

exam (chaired by Prof.\ Sabry) inspires Chapter~\ref{chap:Introduction}.\tabularnewline

@@

\ref{chap:DQT=000026DQC}, and~\ref{chap:QIVPM}.\tabularnewline

@@

is cited in Sec.~\ref{discretequantumtheoryIIb}.\tabularnewline

@@

are cited in Sec.~\ref{modalquantum}.\tabularnewline

@@

cited in Sec.~\ref{subsec:Classical-and-Quantum}.\tabularnewline

@@

\end{doublespace}

@@

To have the current the content of the dissertation, I would also
like to thank John Gardiner for some inspiring discussion about quantum
probability measures over finite fields~\cite{Gardiner2014}. His
results were improved and motivated the transition from DQT to QIVPM
in Sec.~\ref{sec:Toward-IVPM}. Besides of my committee members,
the knowledge I learned from teachers in other courses help me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Although
not directly relating to the dissertation, I have to thank for your
teaching to make me more mature and easier to communicate with others,
especially my English teachers, Traci Nagle, Elizabeth ``Betsy''
Merceron, and Kexin ``Casey'' Chen (谌可心).
\begin{table}[t]
\begin{doublespace}
\noindent \centering{}\caption{\label{tab:teacher}The knowledge I learned from teachers in other
courses help me consolidate this dissertation, where IUB stands for
Indiana University Bloomington and NTU stands for National Taiwan
University.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{Jin-Tzu~Chen}cM{MATH-M522}Y}
\toprule 
Instructor & School & Course & Relation with the dissertation\tabularnewline
\midrule
\midrule 
Tom Lewis & IUB & SLST-T501 & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper of \emph{Academic Writing}.\tabularnewline
\midrule 
Michael\nolinebreak[3] A. Mandell & IUB & MATH-M522 & The textbook~\cite{Hatcher2001} of \emph{Topology II} is cited
in Chapters~\ref{chap:Introduction} and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline
\midrule 
Steven Myers & IUB & CSCI-B502 & The textbook~\cite{AroraBarak2009} of \emph{Computational Complexity}
is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline
\midrule 
Valery Lunts & IUB & MATH-M501 & The textbook~\cite{DummitFoote2004} of \emph{Survey of Algebra}
is cited in Sec.~\ref{sec:background}, \ref{discretequantumtheoryI},
and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline
\midrule 
Huah Chu\linebreak[3] (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & The textbook~\cite{Artin1991} of \emph{Algebra (I)} and \emph{(II)}
is cited in Sec.~\ref{CQC1qubitBloch.sec}, \ref{sec:background},
\ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline
\midrule
Jin-Tzu Chen\linebreak[3] (陳金次) & NTU & 201~31300 & The textbook~\cite{GAMELIN2003} of \emph{Functions of a Complex Variable}
is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{doublespace}
\end{table}

@@

\begin{sidewaystable}
\begin{doublespace}
\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical quantities
and their mathematical models with the inspired quantum models.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YcYY}
\toprule 
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline
\midrule 
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline
\midrule
\midrule 
Mathematical representation & Integer & Real number & Real number\tabularnewline
\midrule 
Computational model & Turing machine & BCSS machine & BCSS machine\tabularnewline
\midrule
\midrule 
How does the theoretical model predict the behavior of the physical
device? & Reliably & Not reliably, but after we have better technology to manipulate the
minimum units, we can more reliably predict the physical phenomenon
and their computational power by a better discrete model. & Not reliably, because precision can never be high enough, and the
difference between their computational power might not be able compensated
by error analysis techniques.\tabularnewline
\midrule 
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{doublespace}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states, which are assumed in the field of complex numbers in the ``conventional
quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}. If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probability'' has well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\begin{comment}
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }
\end{comment}
This idea will be extended to quantum domain as quantum interval-valued
probability measure (QIVPM).

@@

\begin{sidewaystable}
\begin{doublespace}
\noindent \centering{}\caption{\label{tab:organization}Comparison among our models, where the upper
half of the table lists their quantum mechanical definitions and properties,
and the lower half of the table lists their computational power.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YYYYYYY}
\toprule 
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline
\midrule
\midrule 
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline
\midrule 
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline
\midrule 
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline
\midrule 
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline
\midrule
\midrule 
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline
\midrule 
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline
\midrule 
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline
\midrule 
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{doublespace}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and finally, Chapter~\ref{chap:Further-Discussion} contains further
extended questions, especially, briefly explaining why it might be
hard to recover Gleason's theorem for general QIVPMs.

@@

\begin{figure}[!b]
\hfill{}\subfloat[]{\centering{}\includegraphics[width=0.48\columnwidth]{hopfS2}}\hfill{}\subfloat[]{\begin{centering}
\includegraphics[width=0.48\columnwidth]{hopfFibers}
\par\end{centering}
}\hfill{}\caption{\label{hopfFibers.fig}(a) The two-sphere~$\Sphere{2}$ represented
by \Eq{A7.eq}, which is the irreducible space of one-qubit states,
along with a representative set of points on the sphere. Each single
point on the sphere in (a) corresponding to a circle in (b), and a
whole family of circles (the paths of $\rme^{\rmi\theta}$) on the
three-sphere~$\Sphere{3}$ represents the Hopf fibration, Eq.~(\ref{eq:HopfFibration}).
Although $\Sphere{3}$ cannot be directly embedded in $\R^{3}$, three-sphere~$\Sphere{3}$
can be regarded as attaching two three-dimensional balls on two sides
of two-sphere~$\Sphere{2}$. In this way, each circle in $\Sphere{3}$
can be represented as a circle in the three-dimensional ball as shown
in (b). Moreover, points in (a) are color-coded corresponding to circles
in (b), e.g., one pole contains the red elliptical circle that would
become an infinite-radius circle by a slightly different way to represent
$\Sphere{3}$ in $\R^{3}$, and the opposite pole corresponds to the
large perfectly round red circle at the equator.}
\end{figure}
We can reduce 3 degrees of freedom in \Eq{A5.eq} to 2 degrees of
freedom by effectively removing $\rme^{\rmi\theta}$ (``fibering
out by the circle $\Sphere{1}$''). The standard form of these maps
(``the Hopf fibration'') is 
\begin{eqnarray}
 &  & X=2\,\mathrm{Re}\ \alpha_{0}\alpha_{1}^{*}=2x_{0}x_{1}+2y_{0}y_{1}\,,\nonumber \\
 &  & Y=2\,\mathrm{Im}\ \alpha_{0}\alpha_{1}^{*}=2x_{1}y_{0}-2x_{0}y_{1}\,,\label{eq:HopfFibration}\\
 &  & Z=\left|\alpha_{0}\right|^{2}-\left|\alpha_{1}\right|^{2}={x_{0}}^{2}+{y_{0}}^{2}-{x_{1}}^{2}-{y_{1}}^{2}\,.\nonumber 
\end{eqnarray}
By denoting the three-dimensional vector~$\left(X,Y,Z\right)$ as
$\Hat{a}$, \Eq{A5.eq} implies these transformed coordinates obeying
\begin{equation}
\left\Vert \Hat{a}\right\Vert ^{2}=X^{2}+Y^{2}+Z^{2}=\left(\left|\alpha_{0}\right|^{2}+\left|\alpha_{1}\right|^{2}\right)^{2}=1\label{A7.eq}
\end{equation}
and therefore have only two remaining degrees of freedom describing
all possible distinct one-qubit quantum states. In Fig.~\ref{hopfFibers.fig},
we illustrate schematically the family of circles \textit{each one
of which is collapsed to a point}~$\left(\phi,\psi\right)$ on the
surface $X^{2}+Y^{2}+Z^{2}=1$ by the Hopf map.

\begin{figure}[!b]
\begin{centering}
\hfill{}\subfloat[]{\centering{}\includegraphics[width=0.47\columnwidth]{blochSphere}}\hfill{}\subfloat[]{\begin{centering}
\includegraphics[width=0.47\columnwidth]{blochSphereArc}
\par\end{centering}
}\hfill{}
\par\end{centering}
\caption{\label{blochSphere.fig}(a) The conventional Bloch sphere with a unique
state represented by the point at the red sphere. (b) The geodesic
shortest-distance arc connecting two one-qubit quantum states.}
\end{figure}
The resulting manifold is the two-sphere $\Sphere{2}$ (a 2-manifold)
embedded in $\R^{3}$. If we choose one of many possible coordinate
systems describing $\Sphere{3}$ via \Eq{A5.eq} such as 
\begin{equation}
\left(x_{0},y_{0},x_{1},y_{1}\right)=\left(\cos\left(\theta+\phi\right)\cos\psi,\,\sin\left(\theta+\phi\right)\cos\psi,\,\cos\left(\theta-\phi\right)\sin\psi,\,\sin\left(\theta-\phi\right)\sin\psi\right)\,,\label{A8a.eq}
\end{equation}
where %% $0\leq \theta \leq 2 \pi$,  $-\pi \leq \phi \leq  \pi$, 
$0\leq\psi\leq\frac{\pi}{2}$, with $0\leq\theta+\phi<2\pi$ and $0\leq\theta-\phi<2\pi$,
we see that 
\begin{equation}
\left(X,Y,Z\right)=\left(\cos\left(2\phi\right)\sin\left(2\psi\right),\,\sin\left(2\phi\right)\sin\left(2\psi\right),\,\cos\left(2\psi\right)\right)\,.\label{A8b.eq}
\end{equation}
Thus, the one-qubit state is independent of $\theta$, and we can
choose $\theta=\phi$ without loss of generality, reducing the formula
of the unique one-qubit states to $\ket{\psi_{1}}=\rme^{2\rmi\phi}\cos\psi\ket{0}+\sin\psi\ket{1}$,
and an irreducible state can be represented as a point on a sphere
called the Bloch sphere, as shown in Fig.~\ref{blochSphere.fig}~(a).

@@

\begin{figure}[!b]
\noindent \centering{}\includegraphics[width=3.6in]{Fpp2Fp}\caption{\label{proof1.fig}Sketch of the map from $\Fpp$ to $\Fp$ using
$\fnorm{\alpha}$, showing the decomposition of $\Fpp$ into the zero
element $\left(0,0\right)$ and the $p^{2}-1=\left(p+1\right)\left(p-1\right)$
non-zero elements that map onto the $p-1$ non-zero elements of $\Fp$
with multiplicity $p+1$.}
\end{figure}
We can now compute the size of the equivalence class of complex unit-modulus
phases corresponding to the Hopf fibration circle. %
\begin{comment}
\todo{Explain the Hopf fibration circle here\ldots ? }
\end{comment}
Since $\Fpp$ has $p^{2}-1$ non-zero values, and the map $\fnorm{\alpha}$
distributes these equally across the domain of $p-1$ non-zero elements
$c\in\Fp$, there are $\frac{p^{2}-1}{p-1}=p+1$ (non-zero) domain
elements in $\Fpp$ for each (non-zero) image element in $\Fp$. We
illustrate this graphically in Figure~\ref{proof1.fig}. Thus, the
Hopf circle always has the size $p+1$, corresponding essentially
to a discrete projective line, and that is the size of each equivalence
class of the map $\fnorm{\alpha}$ for non-vanishing $\alpha$, including
the map to the unit norm value $c=1\in\Fp$.\end{proof}
